Outline

### 2021-09-08 text/main.Rmd
edits intro, discussion and finnished results

**Main Text:**
The flow of elements through biological communities fuels all ecosystems on Earth [@Welti2017].
Contemporary ecosystems and their functioning are threatened by the harmful dominance of humans.
our modern society has initiated a mass extinction episode, unprecedented in human history and threatening the functioning of the planet’s ecosystems [@Ceballos2015].
Coral reefs represent a prime example of an ecosystem that is severely impacted by anthropogenic threats including warming, and intensive fishing pressure [@Hughes2017; ref fishing].
Coral reefs are among the most diverse and productive ecosystems on Earth and provide essential ecosystem services [@Teh2013].
Severe declines in habitat quality and fish biomass have brought coral reef functioning to the forefront of scientific discourse [@Bellwood2019; @Brandl2019front].
Our capacity to evaluate reef functioning typically relies on static proxies of functions, such as live coral cover, standing stock biomass of reef fishes, or the diversity of qualitative traits [e.g. @Cinner2020].
However, we know comparatively little about actual functions - fluxes of elements and energy - and their drivers (but see [@Allgeier2016]), which constitutes a severe limitation to efficient management [@Brandl2019front]. How can we protect functions if we can not quantify them?

Therefore, trade-offs between functions, their drivers, and their vulnerability remain poorly understood in many ecosystems [@Brandl2019front].

@@

## Results
By combining individual-level bioenergetic models and global underwater visual transect data, we quantified five key ecosystem functions across the globe.
By running a multivariate Bayesian model with, we estimated the average functions per locality as well as a combined measure of multifunction (Fig. 1).
Functions across localities show similar spatial patterns, yet a high level of function does not nescessarily ensure a high level of another function. For example, xxxx.

Biomass is the most commonly employed indicator of coral reef functioning [@Cinner2020; @Brandl2019front], and we observed a predictably strong relationship between fish biomass and all five functions (Fig. 1). We ran a multivariate mixed effect Bayesian model to assess the relationship between all five functions and biomass. We also incorporated sea surface temperature (sst) because it affects the metabolic and groth rates of fishes, which scales up to the community [@Barneche2019].
We found positive slopes xxx for biomass and xxx for sst.

The above-mentioned multivariate model also allowed us to estimate the correlations between functions, independent of the effects of biomass and sst.
Specifically, we estimated correlations between functions on three levels: the locality effects, the site effects, and the residual variations (Fig. 2).
The correlations mostly overlapped across the three levels for a given pair of functions. We found consistently negative trade-offs between P excretion and N excretion, P excretion and biomass production. Further, there were slightly weaker negative correlations between piscivory and N excretion and between piscivory and herbivory. As a consequence of the negative correlations, there are no transects that can have a high value of production and P excretion simultaneously for example.

To explore what drives the residual variation and trade-offs of functions beyond the effects of biomass and sst, we conducted a multivariate Bayesian mixed effect model by adding ten variables that describe the community structure: species richness and the median, lower, and upper 95% quantiles of size, immaturity, and trophic level of the individuals inside a community. We found that the community structure drives the observed trade-offs (Fig. 3).

We also questioned if functions tend to be driven by specific species. We quantified the degree of dominance inside each community and found that functions are on average charachterized by a high level of dominance, i.e. in most communities, a small proportion of species perform most of a function (Fig. 4a). As a comparison, we simulated random communities ranging from 5 to 50 species which yielded an average XX. The predicted average of the degree of dominance was x, x, x, x, and x for each function respectively.
We also calculated the proportion of species that is dominant in at least one transect (i.e. species that contribute more than a community in which each species contributes equally). We found that most species play a dominant role, and 70% of all species contributed disproportionally to a function in at least one community (Fig. 4b).
We finally investigate how often species are dominant across their range and found that very few species are dominant throughout their range (Fig. 4c). Together, this means that functions inside communities tend to be driven by few dominant species, but that the identity of these dominant species varies highly across communities.

@@

By quantifying five key functions mediated by coral reef fishes, we demonstrate that ecosystem-wide functioning is driven by both fish biomass and community structure due to variations in organismal physiology and life-history traits [@Schiettekatte2020; @Schramski2015].
While biomass is one of the most commonly employed indicator of coral reef functioning [@Cinner2020; @Brandl2019front], our analyses suggests that biomass alone does not sufficiently characterize functioning and using biomass as a proxy for functioning masks fundamental differences in critical community-level functions.
Further, we demonstrate strong trade-offs among the five functions, independent of biomass.
Consequently, for a given value of biomass, no reef can yield above average values across all five functions.
While many reefs may stand out as hotspots for one function, none can holistically maximize functioning.

Due to variations in organismal physiology and life-history traits [@Schiettekatte2020], fish community structure significantly impacts ecosystem-wide functioning [@Schramski2015].
For example, phosphorus excretion and piscivory are higher in communities that include large-bodied fishes or occupy high trophic levels (Fig. 2; @Allgeier2014).
In contrast, biomass production is highest in communities dominated by small and/or immature fishes at low trophic levels, creating a trade-off between biomass production and phosphorus excretion.
Metabolic theory predicts that small-bodied individuals have higher mass-specific metabolic rates, leading to elevated consumption rates and disproportionate contributions to functions that rely on rapid energetic turnover [@Barneche2018; @Morais2020glob; @Brandl2019science].
Conversely, fishes in early life stages or with a nutrient-poor diet are often limited by phosphorus [@Schiettekatte2020], resulting in low contributions to phosphorus excretion.

Further, our results reveal that functions consistently hinge on a few dominant species, but the identity of local, dominant species vary strongly across locations [@Lefcheck2019].
Locally, specific high-performing taxa may disproportionately impact rates of functioning at the community level due to high biomass or abundance [@Topor2019].
Specifically, on average, more than 50% of a given function is upheld by only 12% of the species present within a local community.
Despite high species richness on coral reefs, researchers often report the existence of functionally-dominant “key species” [e.g. @Bellwood2006].
While few high-performing taxa dominate functioning in each site, there are no species that are dominant across their entire range.
In addition, 70% of all species contributed disproportionally to a specific function in at least one community.
Our results reveal that while functional dominance is indeed prevalent, the identity of local, dominant species vary strongly across locations [@Lefcheck2019].

Considering multiple functions sheds light on how human-induced shifts in reef fish community structure may impact coral reef ecosystems (Fig. 5).
Anthropogenic stressors have caused severe changes in reef fish biomass and community structure [@Hughes2017], and our findings suggest that these changes will have strong effects on ecosystem functioning.
For example, intensive fishing and associated reductions in biomass of large fishes truncates the size, age, and trophic structure of fish communities [@Graham2017].
When accounting for the effect of biomass, these effects can enhance nitrogen excretion and production [e.g. @Morais2020glob] but negatively impact phosphorus excretion, herbivory, and piscivory (Fig. 2).
On the other hand, declines in coral cover related to climate change are often associated with a shift toward herbivores, which may deter algal domination [@Graham2006].
However, herbivores have a minor contribution to phosphorus excretion [@Allgeier2014; @Schiettekatte2020], so a shift to herbivore dominance and the subsequent decline of community-level phosphorus excretion may change the balance of nutrient cycling on reefs, potentially favoring algal growth [@Burkepile2013].

Similarly, the species-specific vulnerability of functionally-dominant species heavily affects the vulnerability of functions.
By combining species-level vulnerability scores to fishing and climate change induced coral loss [@Graham2011] and the contributions of each species to each function, we illustrate that the loss of individuals most vulnerable to fishing will have greatest impacts on piscivory, followed by phosphorus excretion (Fig. 4).
Conversely, the loss of individuals due to climate change and consequent coral mortality may disproportionally reduce phosphorus excretion, nitrogen excretion, and biomass production.
Combined, fishing and the loss of live coral impact species important for phosphorus excretion.
Surprisingly, although fishing pressure can negatively impact large herbivores such as parrotfishes [@Bellwood2011], herbivory is the least vulnerable function.
This may be due to the high variability in ecosystem roles within the comparatively large pool of herbivorous fish species.
While small herbivores are abundant and not particularly vulnerable to fishing, larger herbivores are frequently targeted and prone to functional extinction in areas with high fishing pressure [@Bellwood2012].
While herbivores of all body sizes and functional groups are combined in our assessment, their realized contributions to herbivory are strongly complementary and, thus, potentially more vulnerable.

Conserving biomass, diversity, and ecosystem functions are important objectives of most conservation initiatives [@Cinner2020].
While safeguarding fish biomass enhances functioning, the trade-offs between key functions reveal a critical challenge for coral reef conservation, where actions to enhance one function may negatively impact another.
For example, the establishment of marine protected areas, which are one of the primary conservation tactics for coral reefs [@Graham2020], may protect herbivorous species and thus provide benefits for herbivory. However, marine protected areas do not protect reefs from the pervasive effects of climate change [@Graham2020], and community shifts towards herbivore domination may result in the decline of phosphorus excretion.
Thus, measuring conservation success with biomass or solely one function (e.g. herbivory) can mask the collapse of other essential functions.
It is necessary to gauge the state of reef ecosystems based on multiple, complementary, process-based functions [@Brandl2019front], as well as making informed decisions on local needs and stressors.
While there is a general consensus on the role of diversity in enhancing functioning [@Zavaleta2010], we highlight the importance of community structure and the identity of dominant species at the local scale.
Maintaining the diversity of fishes is critical, yet, at local scales, species richness has a minor impact on individual functions. Importantly, dissimilarity between local communities may be critical to maintain functioning across seascapes since no species consistently provides high contributions for all functions or across its range [@Zavaleta2010].

@@

![Fig. 1: Spatial variation in five key, biomass-corrected ecosystem functions. Dots indicate locations of field surveys, with dot sizes representing the ranked values of biomass-corrected function, and color scales showing categorical assignments (black = lower 25%, grey = 25-75%, color = >75%). Black circular outlines highlight the five locations with the highest values of each biomass-corrected function.](../output/plots/fig1_multimap.png)

![Fig. 2.  Effects of ecological community variables on five functions. Fixed effect values from Bayesian linear regressions that examine effects of species richness, trophic level, size, and immaturity of fishes. To represent both the median and the spread of trophic level, size, and immaturity across individuals inside a community, we included lower and upper 95% quantile values of these three traits as community variables. All data were log-transformed and standardized to compare across functions and variables (see Table S2 for parameter values on non-standardized data). Dots represent the average effect size estimate, and horizontal lines indicate the 95% credible interval.](../output/plots/fig2_com_slopes.png)

![Fig. 3. Family and species-level contributions to five ecosystem functions on coral reefs. a) The median family-level contributions to each function relative to their contribution to standing stock biomass. The twelve included families are ordered by their median contribution to biomass. b) The distribution of the degree of dominance of communities for each function. Degrees of dominance range between zero (all species contribute equally) and one (a single species is the sole contributor to a given function). c) Species-specific frequencies of dominance in each function across all communities, ranging from zero (species are never dominant) to one (dominant wherever present). A species is categorized as dominant in a community if its contribution to a function is higher than a scenario in which all species are equal (i.e. one divided by the number of species that contribute to the function). Shaded areas show the distribution of the values. Dots represent the median value, and lines indicate the interquartile range.](../output/plots/fig3_contributions.png)

![Fig. 4. Vulnerability of five critical functions to fishing, climate change-induced coral loss, and both stressors combined. Vulnerability is presented as the proportion of communities (filled bars) in which functional vulnerability is higher than vulnerability based on fish biomass (i.e. not accounting for species contributions to each function).](../output/plots/figure_4_vuln_com.png)

>>>>

**Main Text:**
The flow of elements through biological communities fuels all ecosystems on Earth [@Welti2017].
Contemporary ecosystems and their functioning are threatened by the harmful dominance of humans [@Ceballos2015].
Coral reefs represent a prime example of an ecosystem that is severely impacted by anthropogenic threats including warming, and intensive fishing pressure [@Hughes2017; @Pauly2002].
Drastic declines in habitat quality and fish biomass coral reefs have evoked serious concerns about the persistence of coral reef functioning [@Bellwood2019; @Williams2019; @Brandl2019front).
However, our knowledge on process-based functions - fluxes of elements and energy - is still in its infancy and evaluations of coral reef functioning typically relies on static proxies, such as live coral cover, standing stock biomass of reef fishes, or the diversity of qualitative traits [e.g. @Cinner2020].
The sparsity of quantification of functions constitutes a severe limitation to efficient management [@Brandl2019front]: How can we conserve functions if we can not quantify them?

Coral reef fishes play an essential role in coral reef functioning, because coral reef fishes contribute to a high proportion of consumer biomass and therefore represent essential vectors of fluxes of carbon, nitrogen and phosphorus through storage and recycling (McIntyre et al. 2008; All-
geier et al. 2014; Barneche et al. 2014). Broad ecosystem functions mediated by coral reef fishes include nutrient cycling, biomass production, herbivory , and piscivory (secondary consumption) (Brandl et al. 2019front).
While the high diversity of coral reef fishes has inspired many studies with a focus on ecosystem functioning, only a handful of studies have attempted to quantify functions as continuous fluxes (Brandl et al. 2019a). Rather, most studies define fish-mediated ecosystem functioning with proxies such as trait diversity or total biomass (e.g. Mora& Sale 2011; e.g., Mouillot et al. 2014). Further, studies that do quantify functions are mostly focused on biomass production (e.g., @Morais2019; @Morais2020glob) or fish excretion (e.g., @Allgeier2014). An assessment of multiple functions on a large scale is currently lacking.
Therefore, trade-offs between functions, their drivers, and their vulnerability to global stressors remain poorly understood in coral reef ecosystems [@Brandl2019front].

@@

## Results
By combining individual-level bioenergetic models and global underwater visual transect data, we quantified five key ecosystem functions mediated by coral reef fishes across the globe. We estimated the average functions per locality as well as a combined measure of multifunctionality by running a multivariate Bayesian regression model (Fig. 1). Functions across localities show similar spatial patterns, yet a high level of function does not nescessarily ensure a high level of another function. For example, Trindade in the South Atlantic Ocean has high values for N excretion, P excretion, and piscivory, but average values for biomass production. A measure of multifunctionality thus does not appropriately represent the state of all functions.

Biomass is the most commonly employed indicator of coral reef functioning [@Cinner2020; @Brandl2019front], and we demonstrate a predictably strong relationship between fish biomass and all five functions by running multivariate mixed effect Bayesian model (Fig. 1). Specifically, the effects of the log-transformed biomass on log-transformed functions were 0.932 (95\%CI: 0.929; 0.934), 1.051 (1.047; 1.056), 0.771 (0.764; 0.780), 0.940 (0.923; 0.957), and 0.668 (0.635; 0.702) for N excretion, P excretion, production, herbivory, and piscivory, respectively. We also incorporated sea surface temperature (sst) because it affects the metabolic and growth rates of fishes, which scales up to the community [@Barneche2019]. We found positive effects of temperature on N excretion, production, and herbivory, and negligible effects of sst on P excretion and piscivory (suppl table).

The above-mentioned multivariate model also allowed us to estimate the correlations between functions, independent of the effects of biomass and sst.
Specifically, we estimated correlations between functions on three levels: the locality effects, the site effects, and the residual variations (Fig. 2, suppl. fig x).
The correlations mostly overlapped across the three levels for a given pair of functions. We found consistently negative trade-offs between P excretion and N excretion, P excretion and biomass production. Further, there were slightly weaker negative correlations between piscivory and N excretion and between piscivory and herbivory. As a consequence of the negative correlations, a fish community cannot have high values for all functions simultaneously.

To explore what drives the residual variation and trade-offs of functions beyond the effects of biomass and sst, we conducted a multivariate Bayesian mixed effect model by adding ten variables that describe the community structure: species richness and the median, lower, and upper 95% quantiles of size, immaturity, and trophic level of the individuals inside a community. We found that the community structure drives the observed trade-offs (Fig. 3; table x).

We then questioned if functions tend to be driven by specific species across sites. We quantified the degree of dominance inside each community on the site-level and found that functions are on average charachterized by a high level of dominance, i.e. in most communities, a small proportion of species perform most of a function (Fig. 4a). As a comparison, we simulated random communities ranging from 5 to 50 species which yielded an average XX. The predicted average of the degree of dominance was x, x, x, x, and x for each function respectively.
We also calculated the proportion of species that is dominant in at least one transect (i.e. species that contribute more than a community in which each species contributes equally). We found that most species play a dominant role, and 70\% of all species contributed disproportionally to a function in at least one community (Fig. 4b).
We finally investigate how often species are dominant across their range and found that very few species are dominant throughout their range (Fig. 4c). Together, this means that functions inside communities tend to be driven by few dominant species, but that the identity of these dominant species varies highly across communities.

@@

By quantifying five key functions mediated by coral reef fishes, we demonstrate that ecosystem-wide functioning is mainly driven by fish biomass, yet biomass alone does not fully capture functional variability across communities.
While biomass is one of the most commonly employed indicator of coral reef functioning [@Cinner2020; @Brandl2019front], our analyses suggests that biomass alone does not sufficiently characterize functioning and using biomass as a proxy for functioning masks fundamental differences in critical community-level functions.
We demonstrate strong trade-offs among the five functions, independent of biomass.
Consequently, for a given value of biomass, no reef can yield above average values across all five functions.
While many reefs may stand out as hotspots for one function, none can maximize all functions.

Fish community structure drives the observed trade-offs among functions which can be explained by organismal physiology and life-history traits of the individuals inside communities [@Schiettekatte2020; @Schramski2015].
For example, we observed a clear trade-off between P excretion and biomass production which is largely driven by the age and trophic community structure.
Communities dominated by fishes with high trophic levels are charachterized high P excretion rates because predatory fishes have a P-rich diet (@Allgeier2014).
In contrast, biomass production is high in communities dominated by fishes with low trophic levels because herbivores tend to have higher growth rates [@Morais2018].
Moreover, communities dominated by mature fishes have high values of P excretion because growing fishes are more likely to retain P to form P-rich bones [@Schiettekatte2020; ref], while communities dominated by immature fishes logically maximize biomass production.
Furthermore, metabolic theory predicts that small-bodied individuals have higher mass-specific metabolic rates, leading to elevated consumption rates and disproportionate contributions to functions that rely on rapid energetic turnover such as herbivory, piscivory, production, and N excretion [@Barneche2018; @Morais2020glob; @Brandl2019science].

Further, our results reveal that functions consistently hinge on a few dominant species, but the identity of local, dominant species vary strongly across sites [@Lefcheck2019].
Locally, specific high-performing taxa may disproportionately impact rates of functioning at the community level due to high biomass or abundance [@Topor2019].
Specifically, on average, more than 50% of a given function is upheld by only 12% of the species present within a local community.
Despite high species richness on coral reefs, researchers often report the existence of functionally-dominant “key species” [e.g. @Bellwood2006].
While few high-performing taxa indeed dominate functioning in each site, there are no species that are dominant across their entire range.
In addition, more than half of all species contributed disproportionally to a specific function in at least one site inside their range.

Considering multiple functions sheds light on how human-induced shifts in reef fish community structure may impact coral reef ecosystems (Fig. 5).
Fishing and climate-induced coral loss have caused severe changes in reef fish biomass and community structure [@Darling2017; @Graham2017], and our findings suggest that these changes will have strong effects on ecosystem functioning.
Intensive fishing and associated reductions in biomass of large fishes alters the size, age, and trophic structure of fish communities [@Graham2017].
When accounting for the effect of biomass, these effects can enhance nitrogen excretion and production [e.g. @Morais2020glob], but negatively impact phosphorus excretion, herbivory, and piscivory.
Futher, declines in coral cover related to climate change are often associated with a shift toward herbivores, which may deter algal domination [@Graham2006].
However, herbivores have a minor contribution to P excretion [@Allgeier2014; @Schiettekatte2020], so a shift to herbivore dominance and the subsequent decline of community-level P excretion may change the balance of nutrient cycling on reefs, potentially favoring algal growth [@Burkepile2013].

Conserving biomass, diversity, and ecosystem functions are important objectives of most conservation initiatives [@Cinner2020].
While safeguarding fish biomass enhances functioning, the trade-offs between key functions reveal a critical challenge for coral reef conservation, where actions to enhance one function may negatively impact another.
For example, the establishment of marine protected areas, which are one of the primary conservation tactics for coral reefs [@Graham2020], may protect herbivorous species and thus provide benefits for herbivory.
However, marine protected areas do not protect reefs from the pervasive effects of climate change [@Graham2020], and community shifts towards herbivore domination may result in the decline of P excretion.
Thus, measuring conservation success with biomass or solely one function (e.g. herbivory) can mask the collapse of other essential functions.
It is necessary to gauge the state of reef ecosystems based on multiple, complementary, process-based functions, yet our comprehension of
process-based functioning of coral reefs or what a “functional” coral reef really means is still in its infancy [@Brandl2019front].
To appropriately manage coral reef functioning with local needs and stressors, future research should seek to establish functional baselines for global coral reefs.
Until then, our study suggests that coral reef fish functions can be managed by maximizing standing stock biomass, protecting local key species, and promoting a balanced community structure.

@@

![Fig. 1: Maps of five key ecosystem functions and multifunction, and their relationships with biomass. Left: Dots indicate localities of field surveys, with dot sizes representing the ranked values of the locality-level predictions of functions, and color scales showing categorical assignments (black = lower 25\%, grey = 25-75\%, color = >75\%). Black circular outlines highlight the five localities with the highest values of each function. Multifunction represents the weighted average of the five standardized functions. Right: The predicted values for functions and multifunction with increasing biomass.  The lines represent the average predicted value and the shaded areas show the 95% credible interval of he predictions.](../output/plots/fig1_multimap.png)

![Fig. 2: Correlations among five functions while accounting for biomass and sea surface temperature. a: Modeled correlation coefficients of residual errors. Dots represent the average and lines represent the 95%CI. b-k: Scatter plots of the mean residual error of the functions to demonstrate the correlations.](../output/plots/fig1_multimap.png)

![Fig. 3.  Effects of ecological community variables on five functions. Dots indicate fixed effect values from Bayesian linear regressions that examine effects of species richness, trophic level, size, and immaturity of fishes. To represent both the median and the spread of trophic level, size, and immaturity across individuals inside a community, we included lower and upper 95% quantile values of these three traits as community variables. All data were log-transformed and standardized to compare across functions and variables (see Table S2 for parameter values on non-standardized data). Dots represent the average effect size estimate, and horizontal lines indicate the 95% credible interval.](../output/plots/fig2_com_slopes.png)

![Fig. 4: Local dominance in species contributions for five ecosystem functions on coral reefs. a: The degree of dominance of communities for each function on the site-level. Colored dots represent the raw values and the black dots and lines display the mean and 95\% credible intervals of the average degree of dominance, respectively. Degrees of dominance range between zero (all species contribute equally) and one (a single species is the sole contributor to a given function). The vertical dashed line shows the average of the degree of dominance of 1,000 randomly simulated communities. b: Bar plot showing the proportion of species that are dominant in at least one site relative to the total number of species. In the case of herbivory and piscivory, the total number of herbivores and piscivores, respectively, is used.  c: Species-specific frequencies of dominance in each function across all sites, ranging from zero (species are never dominant) to one (dominant wherever present). A species is categorized as dominant in a community if its contribution to a function is higher than a scenario in which all species are equal (i.e. one divided by the number of species that contribute to the function).](../output/plots/fig3_contributions.png)

![Fig. 5. Vulnerability of five critical functions to fishing and climate change-induced coral loss. Conceptual scheme showing the potential ways in which fishing and climate change may affect functions through their effects on biomass and community structure.](../output/plots/figure_4_vuln_com.png)
